---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  cache = TRUE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ritest

<!-- badges: start -->
[![R-CMD-check](https://github.com/grantmcdermott/ritest/workflows/R-CMD-check/badge.svg)](https://github.com/grantmcdermott/ritest/actions)
[![Codecov test coverage](https://codecov.io/gh/grantmcdermott/ritest/branch/master/graph/badge.svg)](https://app.codecov.io/gh/grantmcdermott/ritest?branch=master)
<!-- badges: end -->

Conduct [**randomization inference**](https://dimewiki.worldbank.org/Randomization_Inference)
on R model objects.

This R package is a port of the [`-ritest-`](https://github.com/simonheb) Stata 
routine (by 
[Simon Heß](https://github.com/simonheb)). It doesn't (yet) try to support all
of the features in the Stata version and is currently limited to `lm()` and 
`fixest::feols()` models. But it does appear to be significantly faster, and 
aims to support a variety of model classes once it is fully baked.

[`Installation`](#Installation) | [`Useage`](#Useage) | [`Benchmarks`](#Benchmarks) | [`License`](#)

## Installation

``` r
# install.packages("remotes")
remotes::install_github("grantmcdermott/ritest")
```

## Useage

Here follows a basic example using the `colombia` dataset that comes bundled
with the package. Note that I'll also use the `fixest::feols()` function to 
estimate the parametric model first.

```{r libs, cache=FALSE, message=FALSE}
library(ritest)       ## This package
library(fixest)       ## For fast (high-dimensional) fixed-effect regressions

data("colombia")

## Parametric model using fixest::feols()
co_est = 
  feols(
    dayscorab ~ b_treat + b_dayscorab + miss_b_dayscorab | b_pair + round2 + round3, 
    vcov = ~b_block, data = colombia
    )
```

Let's conduct RI on the treatment variable (`b_treat`), whilst taking account of
the stratified and clustered experimental design.

```{r co_ri}
co_ri = ritest(co_est, 'b_treat', strata='b_pair', cluster='b_block', reps=1e3, seed=546L)
co_ri
```

A default plotting method is provided.

```{r plot_co_ri}
plot(co_ri)
```

For more examples and demonstration of additional features, please see the 
introductory vignette.

## Benchmarks

I generally observe a speed increase of between 25x--50x compared to the Stata 
version. Again, see the introductory vignette for timed examples.

## License

The software code contained within this repository is made available under the 
[MIT license](http://opensource.org/licenses/mit-license.php).
