---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  cache = TRUE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ritest

<!-- badges: start -->
[![R-CMD-check](https://github.com/grantmcdermott/ritest/workflows/R-CMD-check/badge.svg)](https://github.com/grantmcdermott/ritest/actions)
[![Codecov test coverage](https://codecov.io/gh/grantmcdermott/ritest/branch/master/graph/badge.svg)](https://app.codecov.io/gh/grantmcdermott/ritest?branch=master)
<!-- badges: end -->

Conduct [**randomization inference**](https://dimewiki.worldbank.org/Randomization_Inference)
on R model objects.

This R package is a port of the excellent 
[`-ritest-`](https://github.com/simonheb/ritest) Stata routine by Simon Heß. It 
doesn't (yet) try to support all of the features in the Stata version and is 
currently limited to `lm()` and `fixest::feols()` models. But it does appear to 
be significantly faster, and aims to support a variety of model classes once it 
is fully baked.

## Installation

``` r
# install.packages("remotes")
remotes::install_github("grantmcdermott/ritest")
```

## Usage

A detailed walkthrough of the package is provided in the introductory
vignette. See `vignette("ritest")`.

As a quickstart, here follows a basic example using data from a randomized 
control trial (RCT) that was conducted in Colombia. The dataset is provided with 
this package. 

First, we use the `fixest::feols()` function to estimate the parametric model.

```{r co_est, message=FALSE}
library(ritest)  ## This package
library(fixest)  ## For fast (high-dimensional) fixed-effect regressions

data("colombia")

## Parametric model using fixest::feols()
co_est = 
  feols(
    dayscorab ~ b_treat + b_dayscorab | b_pair + miss_b_dayscorab + round2 + round3, 
    vcov = ~b_block, data = colombia
    )
co_est
```

Our key treatment variable (`b_treat`) is deemed to be statistically significant
(p-value of 0.024), even as we cluster the standard errors.

But let's see if this result is robust to randomization inference (RI). We'll 
perform 1,000 RI permutations on `b_treat`, whilst taking into account the 
stratified and clustered experimental design of the underlying RCT.

```{r co_ri}
ritest(co_est, 'b_treat', strata='b_pair', cluster='b_block', reps=1e3, seed=1234)
```

The simulated p-value is noticeably larger than the parametric one (0.107 vs
0.024), suggesting that our parametric model is overstating the effectiveness
of treatment.

For more examples and additional features --- plotting, regression tables, etc. 
--- please see the 
[**introductory vignette**](http://grantmcdermott.com/ritest/articles/ritest.html).

## Benchmarks

I generally observe a speed increase of between 25x--50x compared to the Stata 
version. Again, see the introductory vignette for timed examples.

## Other software

Apart from the Stata [`-ritest-`](https://github.com/simonheb/ritest) routine, 
there are several other packages for conducting randomization inference in R.
For example, the [**ri**](https://cran.r-project.org/web/packages/ri/index.html)
package has been available for nearly a decade. More recently, the successor [**ri2**](https://cran.r-project.org/web/packages/ri2/index.html) package
extends upon the original, with updated syntax and functionality. An advantage
of **ri2** is that it integrates with the wider 
[DeclareDesign](https://declaredesign.org/) suite of R packages for experimental
and empirical research design. This enables researchers to build RI and other
experimental considerations into the incipient design process. On the other
hand, this places some restrictions on conducting RI _ex post_ or in 
quasi-experimental settings (e.g. a study that leverages a natural experiment). 
For example, you can't pass an existing regression model object to 
`ri2::conduct_ri()`, which is what **ritest** was designed for. Your use case 
will likely determine which software is optimal for you.

## License

The software code contained within this repository is made available under the 
[MIT license](http://opensource.org/licenses/mit-license.php).
