---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  cache = TRUE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ritest

<!-- badges: start -->
[![R-CMD-check](https://github.com/grantmcdermott/ritest/workflows/R-CMD-check/badge.svg)](https://github.com/grantmcdermott/ritest/actions)
<!-- badges: end -->

An experimental R port of the **[`ritest`](https://github.com/simonheb)** Stata 
routine for randomization inference (RI) by 
[Simon Heß](https://github.com/simonheb). 

Fast and user-friendly. Currently limited to `lm()` and `fixest::feols()` 
models, but aims to support a variety of model classes once it is fully baked.

## Installation

``` r
# install.packages("remotes")
remotes::install_github("grantmcdermott/ritest")
```

## Examples

Let's start by loading the **ritest** package. 
I'll also load the **fixest** and **haven** packages to help demonstrate some
additional functionality in the examples that follow.

```{r libs, cache=FALSE, message=FALSE}
library(ritest)
library(fixest) ## Note: ritest requires fixest version >= 0.10.0
library(haven)  ## For reading .dta files
```

The `ritest()` function supports a variety of arguments, but the basic syntax 
is

```r
ritest(object, resampvar, reps=100, strata=NULL, cluster=NULL, ...)
```

where: 

- `object` is a model object (e.g. a linear regression).
- `resampvar` is the variable that you want to resample (i.e. permute).
- `reps` is the number of simulations (i.e. permutations).
- `strata` is a variable defining the stratification (aka "blocking") of the 
experimental design, if any.
- `cluster` is a variable defining the clustering of treatment in the 
experimental design, if any.
- `...` are other arguments. This includes the ability to set a random seed for
reproducibility, controlling the parallelism behaviour, adding a progress bar,
etc. See `?ritest` for more information.


### Toy data

Our first example will be a rather naive implementation using the inbuilt
[`nkp`](https://vincentarelbundock.github.io/Rdatasets/doc/MASS/npk.html) 
dataset.

```{r est}
est = lm(yield ~ N + P + K, data = npk)
```

Let's say we're interested in the yield effect of 'N' (i.e. nitrogen).
We want to know whether our inferential reasoning about this parameter is 
robust to using RI, as opposed to just relying on the parametric t-test and 
p-value produced by our linear regression model. We'll do 1,000
simulations and, just for illustration, limit the number of parallel cores to 2. 
(The default parallel behaviour will use half of the available cores on a user's
machine.) 
The 'verbose = TRUE' argument simply prints the results upon completion,
including the original regression model summary.

```{r est_ri}
# est_ri = ritest(est, ~N, reps = 1e3, seed = 1234L, pcores = 2L, verbose = TRUE) ## Formulas work too
est_ri = ritest(est, 'N', reps = 1e3, seed = 1234L, pcores = 2L, verbose = TRUE)
```

In this simple case, our parametric results appear to hold up very well. The
original p-value of 0.019 is very close to the equivalent rejection rate
of 0.021 that we get with our RI procedure.

We can also visualize this result using the default `plot` method. The function 
takes several arguments for added customization. But here I'll just show how to
add the parametric 95 percent confidence interval (i.e. from the original linear 
regression) in addition the simulated rejection regions.

```{r plot_est_ri}
plot(est_ri, show_parm = TRUE)
```

A more realistic implementation would control for the stratified (aka "blocked")
experimental design. We'll specify these strata (blocks) as fixed-effects in 
our regression model. I'll also use this as an opportunity to show that that the
package is fully compatible with piping workflows. This might be useful if you
don't feel like saving intermediate objects. Here I'll use the new base R pipe 
(`|>`) that was introduced in R 4.1.1, but the same thing would be possible with
the popular magrittr pipe (`%>%`).

```{r est2}
# library(fixest) ## Already loaded

feols(yield ~ N + P + K | block, vcov = 'iid', data = npk) |>
  ritest('N', strata = 'block', reps = 1e3, seed = 99L)
```

Again, we obtain an RI rejection rate (0.003) that is very 
similar to the parametric p-value (0.004) from the regression model.

### Real-life data

Next, we'll replicate an example in David McKenzie's nice 
[blog post](https://blogs.worldbank.org/impactevaluations/finally-way-do-easy-randomization-inference-stata)
about the Stata `ritest` routine. The dataset in question derives from a 
randomized control trial about supply chains in Columbia. 
You can download the data from David's website 
[here](https://drive.google.com/open?id=0B9C9RwWKZrUNazdyVXFkSTlTNGc) (also 
includes code and data for some other examples.)

#### Stata implementation

First, here is the Stata code and output. I won't go into detail about the 
actual research question that we're trying to address. (See David's blog post 
for more on that.) The essential thing to know is that I'm going to run 
5,000 RI permutations on a pretty standard fixed-effect model, whilst 
accounting for the stratified and clustered design of the experiment. 

(Aside: I'm also snipping most of the Stata output, so as to only highlight the 
main command and result.)

```stata
. use "~/clusterColombia.dta", clear

. 
. timer on 1

. 
. ritest b_treat _b[b_treat], cluster(b_block) strata(b_pair) reps(5e3) seed(546): ///
> areg dayscorab b_treat b_dayscorab miss_b_dayscorab round2 round3, cluster(b_block) a(b_pair)

[snipped]


      command:  areg dayscorab b_treat b_dayscorab miss_b_dayscorab round2 round3, cluster(b_block)
                    a(b_pair)
        _pm_1:  _b[b_treat]
  res. var(s):  b_treat
   Resampling:  Permuting b_treat
Clust. var(s):  b_block
     Clusters:  63
Strata var(s):  b_pair
       Strata:  31

------------------------------------------------------------------------------
T            |     T(obs)       c       n   p=c/n   SE(p) [95% Conf. Interval]
-------------+----------------------------------------------------------------
       _pm_1 |   -.180738     529    5000  0.1058  0.0043  .0974064   .1146569
------------------------------------------------------------------------------
Note: Confidence interval is with respect to p=c/n.
Note: c = #{|T| >= |T(obs)|}

. 
. timer off 1

```

Like David, this takes around **3 minutes** to run on my laptop.

```stata
. timer list
   1:    183.01 /        1 =     183.0150
```

#### R implementation

Now, the R equivalent with this package.

```{r co_ri}
co = haven::read_dta('~/clusterColombia.dta')

co_est = 
  feols(
    dayscorab ~ b_treat + b_dayscorab + miss_b_dayscorab + round2 + round3 | b_pair, 
    vcov = ~b_block, data = co
    )

tic = Sys.time()
co_ri = ritest(co_est, ~b_treat, cluster=~b_block, strata=~b_pair, reps=5e3, seed=546L)
toc = Sys.time() - tic

## Print the results
co_ri
```

Using the same random seed is a bit of performance art. We won't get exactly
the same results across two different languages. But the important thing to note 
is that they are functionally equivalent (rejection probability of 0.106 vs
0.104).

One nice feature of this implementation is that it should complete very quickly. 
This time, the 5,000 simulations only take around **6 seconds**.

```{r toc}
toc
```

Again, we can plot the results. Here's a slight variation, where we plot in
histogram form and use a fill to highlight the 95% rejection region(s) instead
of vertical lines.

```{r plot_co_ri}
plot(co_ri, type = 'hist', highlight = 'fill')
```


### Benchmarks

As indicated in the real-life data example, I generally observe a speed increase
of around 30x compared to the Stata version. This gap appears to widen further
for larger datasets and for certain kinds of RI setups.
