---
title: "Introduction to ritest"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ritest}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  cache = TRUE,
  collapse = TRUE,
  comment = "#>"
)
```

Let's start by loading the **ritest** package. I'll also be using one or two 
outside packages in the examples that follow to demonstrate some additional 
functionality, but will hold off loading those for now.

```{r libs, cache=FALSE, message=FALSE}
library(ritest)
```

The `ritest()` function supports a variety of arguments, but the basic syntax 
is

```r
ritest(object, resampvar, reps=100, strata=NULL, cluster=NULL, ...)
```

where: 

- `object` is a model object (e.g. a linear regression).
- `resampvar` is the variable that you want to resample (i.e. permute). You can
also specify the sharp null hypothesis that you want to test as part of a 
character string.
- `reps` is the number of simulations (i.e. permutations).
- `strata` is a variable defining the stratification (aka "blocking") of the 
experimental design, if any.
- `cluster` is a variable defining the clustering of treatment in the 
experimental design, if any.
- `...` are other arguments. This includes the ability to set a random seed for
reproducibility, controlling the parallelism behaviour, adding a progress bar,
etc. See `?ritest` for more information.


Let's see this functionality in action with the help of some examples.

## Example I: Toy data

Our first example will be a rather naive implementation using the base
[`nkp`](https://vincentarelbundock.github.io/Rdatasets/doc/MASS/npk.html) 
dataset.

```{r est}
est = lm(yield ~ N + P + K, data = npk)
```

Let's say we're interested in the yield effect of 'N' (i.e. nitrogen).
We want to know whether our inferential reasoning about this parameter is 
robust to using RI, as opposed to just relying on the parametric t-test and 
p-value produced by our linear regression model. We'll do 1,000
simulations and, just for illustration, limit the number of parallel cores to 2. 
(The default parallel behaviour will use half of the available cores on a user's
machine.) 
The 'verbose = TRUE' argument simply prints the results upon completion,
including the original regression model summary.

```{r est_ri}
est_ri = ritest(est, 'N', reps = 1e3, seed = 1234L, pcores = 2L, verbose = TRUE)
```

In this simple case, our parametric results appear to hold up very well. The
original p-value of 0.019 is very close to the equivalent rejection rate
of 0.021 that we get with our RI procedure.

We can also visualize this result using the dedicated `plot` method. The function 
takes several arguments for added customization. But here I'll just show the 
default plot, which includes vertical lines that denote the simulated (in this 
case: 95 percent) rejection regions.

```{r plot_est_ri}
plot(est_ri)
```

As an aside, note that the RI procedure tests against a standard two-sided null 
hypothesis of zero. (In the above case: `H0: N1=0`.) We can specify a different
null hypothesis as part of the `resampvar` string. For example:

```{r plot_est_ri_alt_ho}
plot(ritest(est, 'N<=2', reps = 1e3, seed = 1234L, pcores = 2L))
```

Note that we could (and probably should) have estimated a more realistic model
that controls for the stratified (aka "blocked") design of the original `npk`
experiment. This is easily done, but we'll hold off doing so for now since the
the discussion of strata provides a nice segue to our next example.

## Example II: Real-life data

Our second example will provide a more realistic use-case, where we need to
account for a _stratified_ and _clustered_ research design. 
In particular, we'll replicate a real-life experiment that David McKenzie
describes in a very helpful
[blog post](https://blogs.worldbank.org/impactevaluations/finally-way-do-easy-randomization-inference-stata) on the original Stata `-ritest-` routine. 

The dataset in question derives from a randomized control trial about supply 
chains in Colombia, which David has kindly provided for re-use in this package 
(see: `?colombia`). The key research question that we're trying to answer below 
is whether a treatment intervention (`b_treat`) led to a drop in the number of 
days requiring visits to the _Corabastos_ central 
market.^[Repeated visits to the market are an expensive and time-consuming exercise for the fresh produce vendors that formed the study population.] 
Moreover, we want to know if our inference about this treatment effect is robust 
to RI.

### Stata implementation

As a benchmark, first we recapitulate David's Stata code and output. I won't go 
into details --- the essential thing to know is that I'm going to run 5,000 RI 
permutations on a pretty standard fixed-effect model, whilst 
accounting for the stratified and clustered design of the experiment. 

(Aside: I'm also snipping most of the Stata output, so as to only highlight the 
main command and result.)

```stata
. // This next line assumes you have exported the `colombia` dataset from R as a
. // CSV for Stata to read, e.g. `write.csv(colombia, '~/colombia.csv', row.names = FALSE)`
. insheet using "~/colombia.csv", comma clear

. 
. timer on 1

. 
. ritest b_treat _b[b_treat], cluster(b_block) strata(b_pair) reps(5e3) seed(546): ///
> areg dayscorab b_treat b_dayscorab miss_b_dayscorab round2 round3, cluster(b_block) a(b_pair)

[snipped]


      command:  areg dayscorab b_treat b_dayscorab miss_b_dayscorab round2 round3, cluster(b_block)
                    a(b_pair)
        _pm_1:  _b[b_treat]
  res. var(s):  b_treat
   Resampling:  Permuting b_treat
Clust. var(s):  b_block
     Clusters:  63
Strata var(s):  b_pair
       Strata:  31

------------------------------------------------------------------------------
T            |     T(obs)       c       n   p=c/n   SE(p) [95% Conf. Interval]
-------------+----------------------------------------------------------------
       _pm_1 |   -.180738     529    5000  0.1058  0.0043  .0974064   .1146569
------------------------------------------------------------------------------
Note: Confidence interval is with respect to p=c/n.
Note: c = #{|T| >= |T(obs)|}

. 
. timer off 1

```

Like David, this takes around **3 minutes** to run on my laptop.

```stata
. timer list
   1:    183.01 /        1 =     183.0150
```

### R implementation

Let's replicate the above in R using this package. First, we'll estimate and 
save the parametric model using `fixest::feols()`.

```{r co_est}
data("colombia")

library(fixest) ## For fast (high-dimensional) fixed-effects models

co_est = 
  feols(
    dayscorab ~ b_treat + b_dayscorab + miss_b_dayscorab | b_pair + round2 + round3, 
    vcov = ~b_block, data = colombia
    )
co_est
```

Now, we conduct RI on our model to see whether our key treatment variable 
(`b_treat`) is sensitive to the imposed parametric constraints. Note that we
can specify the strata and clusters as additional arguments to `ritest()`. 

```{r co_ri}
tic = Sys.time() ## timer on

co_ri = ritest(co_est, 'b_treat', cluster='b_block', strata='b_pair', reps=5e3, seed=546L)

toc = Sys.time() - tic ## timer off

## Print the results
co_ri
```

Using the same random seed in R and Stata is a bit of performance art. We won't
get exactly
the same results across two different languages. But the important thing to note 
is that they are functionally equivalent (rejection probability of 0.106 vs
0.104). More importantly, we can see that our inference about the effectiveness
of treatment in this study is indeed sensitive to RI. Our parametric p-value 
(0.024) is much lower than the permuted rejection rate (0.104).

Again, we can plot the results. Here's a slight variation, where we plot in
histogram form and use a fill to highlight the 95% rejection region(s) instead
of vertical lines.

```{r plot_co_ri}
plot(co_ri, type = 'hist', highlight = 'fill')
```

### Benchmarks

One nice feature of the R implementation is that it should complete very 
quickly. Instead of taking 3 minutes, this time the 5,000 simulations only take
around **6 seconds**.

```{r toc, eval=FALSE}
toc
#> Time difference of 6.581697 secs
```

As a general observation, the R implementation of `ritest()` doesn't yet offer 
all of the functionality of the Stata version. For example, it doesn't support
an external file of resampling weights. However, it does appear to be a lot
(25x -- 50x) faster and this might make it more suitable for certain types of
problems.

## Extras and asides

### Regression tables

Support for regression tables is enabled via **ritest's** compatability
with the **modelsummary** package. I recommend displaying p-values instead of 
the default standard errors. This is particularly important when comparing 
against a parametric model, as I do below.

```{r msummary_co_ri}
library(modelsummary)

msummary(list(lm = co_est, ritest = co_ri), 
         statistic = 'p.value', 
         ## These next arguments just make our comparison table look a bit nicer
         coef_map = c('b_treat' = 'Treatment'), 
         gof_omit = 'Obs|R2|IC|Log|F',
         notes = 'p-values shown in parentheses.')
```

### Formulas

Formula interfaces are supported if you don't like writing variables (i.e.
the `resampvar` and/or `strata` and `cluster` arguments) as strings. I'll just
use the default number of reps (i.e. 100) and drop the random seed for this next
example.

```{r formulas}
ritest(co_est, ~b_treat, strata=~b_pair, cluster=~b_block)
```

### ggplot2

If you don't like the default plot method and would prefer to use **ggplot2**,
then that's easily done. Just extract the beta values from the return object.

```{r ggplot2_co_ri}
library(ggplot2)

ggplot(data.frame(betas = co_ri$betas), aes(betas)) + 
  geom_density() +
  theme_minimal()
```

### Piping workflows

The **ritest** package is fully compatible with piping workflows. This might be
useful if you don't feel like saving intermediate objects. Here I'll demonstrate 
using the new base R pipe (`|>`) that was introduced in R 4.1.0, but the same 
principal would carry over to popular magrittr pipe (`%>%`).

```{r est2_pipe}
## Only run this code if you have R 4.1 or above
if (getRversion() >= "4.1") {
  feols(yield ~ N + P + K | block, vcov = 'iid', data = npk) |>
  ritest('N', strata = 'block', reps = 1e3, seed = 99L) |>
  plot()
}

```
